Openness and Computational Reproducibility in plant pathology: where do we stand and a way forward

### 2022-06-04 analysis/paper/paper.Rmd
Remove trailing space
Clarify discussion

At the second level, in our search of the literature, we found that @Hill2019 shared code and example data to reproduce the results of the paper as supplementary materials, citing all packages used and the versions.
The scripts provided used an R package, *checkpoint* [@Ooi2022], to provide a mechanism to help ensure reproducibility by installing the package versions used by the authors that are necessary to reproduce the work rather than defaulting to the latest versions.
However, the scripts did not run unhindered, as the supplementary materials suggested that they should, without changes to the data, but with some modifications to the data, and file name changes, the scripts ran allowing examples of the research to be reproduced using the definition of @Peng2009.

The third level is the use of proper code repositories such as GitHub and data repositories like FigShare, Zenodo or OSF.io among others, allow for the deposition and updating of code, figures, data preprints or any other materials that support the article itself while providing a DOI and citation for a standalone citation.
Sparks et al. [@Sparks2011; @Sparks2014] used FigShare to provide models, data and code, [@Sparks2016] necessary to replicate model development and the subsequent study on the effects of climate change on potato late blight.
Similarly, @carleson2019population hosted the code for reproducing a population genomic analysis of *Phytophthora plurivora* on GitHub, while providing all data on OSF.io (the Open Science Framework).
@Lehner2017 used GitHub to host a code repository of their research compendium website with data and a reproducible report that explains in detail all steps of the analysis and the R code for conducting a meta-analysis for assessing heterogeneity in the relationship between white mold incidence and soybean yield and between incidence and soybean yield (<https://emdelponte.github.io/paper-white-mold-meta-analysis/>).
The website clearly demonstrates the analysis to readers and uses R so that anyone can easily replicate the work.
Likewise, @Duku2016 provide scripts for models, data and code for graphs, via a website hosted by GitHub, <http://adamhsparks.github.io/MICCORDEA/>, that can be used to replicate their study modelling the effects of climate change on rice bacterial blight and rice leaf blast in Tanzania.
However, even this example is not totally open as it requires ArcGIS, an expensive proprietary software, to run the provided Python scripts for the RICEPEST model [@Savary2014].
This highlights the difficulty in making work completely open and reproducible, further work could be undertaken to make the Python scripts rely totally on free-open-source-software but the work as it it provided here does allow for the complete reproducibility of the paper as written provided with the proper tools.

>>>>

At the second level, in our search of the literature, we found that @Hill2019 shared code and example data to reproduce the results of the paper as supplementary materials, citing all packages used and the versions.
The scripts provided used an R package, *checkpoint* [@Ooi2022], to provide a mechanism to help ensure reproducibility by installing the package versions used by the authors that are necessary to reproduce the work rather than defaulting to the latest versions.
However, the scripts did not run unhindered, as the supplementary materials suggested that they should, without changes to the data, but with some modifications to the data, and file name changes, the scripts ran allowing examples of the research to be reproduced using the definition of @Peng2009.
One drawback to this approach is that journals are often not equipped to handle code, e.g., script files, that may be developed as a part of the research process and so they are often archived as PDF or Word documents which hinder the ability to easily ingest and start working with them.
Though this should be possible given some effort from the publishing journal to share simple text files rather than binary formats along with the proper instructions and handling of these file formats rather than a one-size-fits all approach that we commonly see for supplementary materials.

The third level is the use of proper code repositories such as GitHub and data repositories like FigShare, Zenodo or OSF.io among others, allow for the deposition and updating of code, figures, data preprints or any other materials that support the article itself while providing a DOI and citation for a standalone citation.
As an example, Sparks et al. [-@Sparks2011; -@Sparks2014] used FigShare to provide models, data and code [@Sparks2016] necessary to replicate model development and the subsequent study on the effects of climate change on potato late blight.
Similarly, @carleson2019population hosted the code for reproducing a population genomic analysis of *Phytophthora plurivora* on GitHub, while providing all data on OSF.io (the Open Science Framework).
@Lehner2017 used GitHub to host a code repository of their research compendium website with data and a reproducible report that explains in detail all steps of the analysis and the R code for conducting a meta-analysis for assessing heterogeneity in the relationship between white mold incidence and soybean yield and between incidence and soybean yield (<https://emdelponte.github.io/paper-white-mold-meta-analysis/>).
The website clearly demonstrates the analysis to readers and uses R so that anyone can easily replicate the work.
Likewise, @Duku2016 provide scripts for models, data and code for graphs, via a website hosted by GitHub, <http://adamhsparks.github.io/MICCORDEA/>, that can be used to replicate their study modelling the effects of climate change on rice bacterial blight and rice leaf blast in Tanzania [@Sparks2022a].
However, even this example is not totally open as it requires ArcGIS, an expensive proprietary software, to run the provided Python scripts for the RICEPEST model [@Savary2014].
This highlights the difficulty in making work completely open and reproducible, further work could be undertaken to make the Python scripts rely totally on free-open-source-software but the work as it it provided here does allow for the complete reproducibility of the paper as written provided with the proper tools.
Using a repository resource like GitHub allows readers to easily contact the authors by opening "Issues" and report bugs or ask questions in an open forum that are not as straightforward when the data are provided as supplementary material.
